Load libraries

```{r}
library(sf) 
library(here)
library(tidyverse)
library(tmap)
library(tmaptools)
library(raster) 
library(terra)
library(fs)
library(ggplot2)
```
Read Vietnam outline shapefile

```{r}
vietnamoutline <- st_read(here(
  "data",
  "vnm_adm_gov_20201027",
  "vnm_admbnda_adm0_gov_20200103.shp"
))
st_crs(vietnamoutline)$proj4string # Check projection
plot(vietnamoutline) # Plot
```
Read raster file

```{r}
future21_40 <-
  terra::rast(
    here("data", "wc2.1_2.5m_tmax_ACCESS-CM2_ssp370_2021-2040.tif")
  )

future41_60 <-
  terra::rast(
    here("data", "wc2.1_2.5m_tmax_ACCESS-CM2_ssp370_2041-2060.tif")
  )

future61_80 <-
  terra::rast(
    here("data", "wc2.1_2.5m_tmax_ACCESS-CM2_ssp370_2061-2080.tif")
  )

future81_00 <-
  terra::rast(
    here("data", "wc2.1_2.5m_tmax_ACCESS-CM2_ssp370_2081-2100.tif")
  )
```
Rename each layer into month
```{r}
month <-
  c("Jan",
    "Feb",
    "Mar",
    "Apr",
    "May",
    "Jun",
    "Jul",
    "Aug",
    "Sep",
    "Oct",
    "Nov",
    "Dec")

names(future21_40) <- month
names(future41_60) <- month
names(future61_80) <- month
names(future81_00) <- month
```

Use a shape instead of point to carve out data from raster
```{r}
vietnamoutline %>% 
  st_simplify(.,dTolerance = 5000) %>% 
  st_geometry()%>% 
  plot()

crs(vietnamoutline)
```
Crop and mask
```{r}

vntemp21_40 <- future21_40 %>% 
  terra::crop(.,vietnamoutline) %>% 
  terra::mask(.,vietnamoutline)
vntemp41_60 <- future41_60 %>% 
  terra::crop(.,vietnamoutline) %>% 
  terra::mask(.,vietnamoutline)
vntemp61_80 <- future61_80 %>% 
  terra::crop(.,vietnamoutline) %>% 
  terra::mask(.,vietnamoutline)
vntemp81_00 <- future81_00 %>% 
  terra::crop(.,vietnamoutline) %>% 
  terra::mask(.,vietnamoutline)
```
Convert to tibble and dataframes

```{r}
vntemp_tib <- vntemp %>% 
  as_tibble()

vntemp_df <- vntemp %>% 
  as.data.frame()
```

Extract July
```{r}
jul2140 <- vntemp21_40[[7]]
jul4160 <- vntemp41_60[[7]]
jul6180 <- vntemp61_80[[7]]
jul8100 <- vntemp81_00[[7]]

plot(
  jul2140,
  col= rev(heat.colors(999)),
  axes = FALSE,
  main = "Temperature forecast in July 2021-2040",
  cex.main = 0.8,
  range = c(20, 50),
  plg = list(shrink = 0.8,
             title = "째C")
)
plot(
  jul4160,
  col= rev(heat.colors(999)),
  axes = FALSE,
  main = "Temperature forecast in July 2041-2060",
  cex.main = 0.8,
  range = c(20, 50),
  plg = list(shrink = 0.8,
             title = "째C")
)
plot(
  jul6180,
  col= rev(heat.colors(999)),
  axes = FALSE,
  main = "Temperature forecast in July 2061-2080",
  cex.main = 0.8,
  range = c(20, 50),
  plg = list(shrink = 0.8,
             title = "째C")
)
plot(
  jul8100,
  col= rev(heat.colors(999)),
  axes = FALSE,
  main = "Temperature forecast in July 2081-2100",
  cex.main = 0.8,
  range = c(20, 50),
  plg = list(shrink = 0.8,
             title = "째C")
)

```





















# Pivot longer so that we have multiple months on the histogram and not just one

squishdata\<-exactAusdf%\>% pivot_longer( cols = 1:12, names_to = "Month", values_to = "Temp" )

# Subsetting two months

twomonths \<- squishdata %\>% filter(., Month=="Jan" \| Month=="Jun")

# mean two months

meantwomonths \<- twomonths %\>% group_by(Month) %\>% summarise(mean=mean(Temp, na.rm=TRUE))

meantwomonths

# PLot them nicely!

ggplot(twomonths, aes(x=Temp, color=Month, fill=Month)) + geom_histogram(position="identity", alpha=0.5)+ geom_vline(data=meantwomonths, aes(xintercept=mean, color=Month), linetype="dashed")+ labs(title="Ggplot2 histogram of Australian Jan and Jun temperatures", x="Temperature", y="Frequency")+ theme_classic()+ theme(plot.title = element_text(hjust = 0.5))

# Plot all 12 months!

data_complete_cases \<- squishdata %\>% drop_na()%\>% mutate(Month = factor(Month, levels = c("Jan","Feb","Mar", "Apr","May","Jun", "Jul","Aug","Sep", "Oct","Nov","Dec"))) view(data_complete_cases)

# Plot faceted histogram

ggplot(data_complete_cases, aes(x=Temp, na.rm=TRUE))+ geom_histogram(color="black", binwidth = 5)+ labs(title="Ggplot2 faceted histogram of Australian temperatures", x="Temperature", y="Frequency")+ facet_grid(Month \~ .)+ theme(plot.title = element_text(hjust = 0.5))

# Using plotly

library(plotly)

jan \<- squishdata %\>% drop_na() %\>% filter(., Month=="Jan")

jun \<- squishdata %\>% drop_na() %\>% filter(., Month=="Jun")

# give axis titles

x \<- list (title = "Temperature") y \<- list (title = "Frequency")

# set the bin width

xbinsno\<-list(start=0, end=40, size = 2.5)

# plot the histogram calling all the variables we just set

ihist \<- plot_ly(alpha = 0.6) %\>% add_histogram(x = jan$Temp,  xbins=xbinsno, name="January") %>%  add_histogram(x = jun$Temp, xbins=xbinsno, name="June") %\>% layout(barmode = "overlay", xaxis=x, yaxis=y)

ihist

# mean per month

meanofall \<- squishdata %\>% group_by(Month) %\>% summarise(mean = mean(Temp, na.rm=TRUE))

# print the top 1

head(meanofall, n=1)

# standard deviation per month

sdofall \<- squishdata %\>% group_by(Month) %\>% summarize(sd = sd(Temp, na.rm=TRUE))

# maximum per month

maxofall \<- squishdata %\>% group_by(Month) %\>% summarize(max = max(Temp, na.rm=TRUE))

# minimum per month

minofall \<- squishdata %\>% group_by(Month) %\>% summarize(min = min(Temp, na.rm=TRUE))

# Interquartlie range per month

IQRofall \<- squishdata %\>% group_by(Month) %\>% summarize(IQR = IQR(Temp, na.rm=TRUE))

# perhaps you want to store multiple outputs in one list..

lotsofstats \<- squishdata %\>% group_by(Month) %\>% summarize(mean = mean(Temp, na.rm=TRUE), max=max(Temp, na.rm=T), min = min(Temp, na.rm=TRUE), sd = sd(Temp, na.rm=TRUE), IQR = IQR(Temp, na.rm=TRUE))

view(lotsofstats)

# or you want to know the mean (or some other stat)

# for the whole year as opposed to each month...

meanwholeyear \<- squishdata %\>% summarize(mean = mean(Temp, na.rm=TRUE), max=max(Temp, na.rm=T), min = min(Temp, na.rm=TRUE), sd = sd(Temp, na.rm=TRUE), IQR = IQR(Temp, na.rm=TRUE))

view(meanwholeyear)

\`\`\`

Using a raster stack we can extract data with a single command!! For example let's make a dataframe of some sample sites --- Vietnam cities/towns

```{r}
worldcities <- st_read(
  here("..", "gis",
       "data",
       "World_Cities.geojson")
)

vn_cities <- worldcities %>%
  dplyr::filter(., .$CNTRY_NAME == "Vietnam") %>%
  dplyr::select(., c("CITY_NAME", "geometry"))
vn_cities
```
Extract the data from the Rasterstack for all points

```{r}
vncitytemp <- terra::extract(future21_40, vn_cities) %>%
  as_tibble() %>%
  add_column(Site = vn_cities$CITY_NAME, .before = "Jan")
View(vncitytemp)
```