---
editor_options: 
  markdown: 
    wrap: 72
---

**The task this week is to:**

-   Read in [global gender inequality
    data](https://hdr.undp.org/data-center/documentation-and-downloads)

-   Join the global gender inequality index to [spatial data of the
    World](https://hub.arcgis.com/datasets/2b93b06dc0dc4e809d3c8db5cb96ba69_0/explore?location=-2.688200%2C0.000000%2C1.41),
    creating a new column of difference in inequality between 2010 and
    2019

-   Share it with the World on GitHub

-   Add you repository URL to the circulated spreadsheet

[**Tip:**]{.underline}

-   The [countrycode](https://github.com/vincentarelbundock/countrycode)
    R package will be helpful!

-   The gender inequality has changed in the last year, you will find
    what you need in the "All composite indices and components time
    series (1990-2021)" dataset, the metadata file beneath it will
    explain what the columns are.

------------------------------------------------------------------------

### Install packages and load them in

```{r include=FALSE}

# install packages
install.packages('remotes')
#load packages
library(sf)
library(tidyverse)
library(janitor)
library(terra)
library(here)
library(remotes)
install_github('vincentarelbundock/countrycode')
library(countrycode)
library(tmap)
library(sf)
```

------------------------------------------------------------------------

### Read the necessary data

```{r}
#Read hdr data and manipulate to produce gii_diff_c
hdrdata <-
  read_csv(here("data",
    "HDR21-22_Composite_indices_complete_time_series.csv"
  )) %>%
  dplyr::select(c("iso3", "country", "region", "gii_2010", "gii_2019","gdi_2010", "gdi_2019","hdi_2010", "hdi_2019"))

# change column type as numeric

hdrdata$gii_2010 <- as.numeric(hdrdata$gii_2010)
hdrdata$gii_2019 <- as.numeric(hdrdata$gii_2019)
hdrdata$gdi_2010 <- as.numeric(hdrdata$gdi_2010)
hdrdata$gdi_2019 <- as.numeric(hdrdata$gdi_2019)
hdrdata$hdi_2010 <- as.numeric(hdrdata$hdi_2010)
hdrdata$hdi_2019 <- as.numeric(hdrdata$hdi_2019)
```

```{r}
View(world)
```

```{r}
#Read world map & join it with hdr data
world <-
  st_read(here("data", "World_Countries_Generalized.geojson")) %>%
  dplyr::mutate(iso3 = countrycode(ISO, origin = 'iso2c', destination =
                                     'iso3c'),
                .before = ISO) %>%
  dplyr::left_join(., hdrdata, by = 'iso3') %>%
  dplyr::mutate(gii_diff = gii_2019 - gii_2010) %>%
  dplyr::mutate(
    gii_diff_c = case_when(
      gii_diff > 0 ~ "More unequal",
      gii_diff == 0 ~ "No change",
      gii_diff < 0 ~ "More equal",
      is.na(gii_diff) ~ "No Data"
    )
  ) %>%
  dplyr::mutate(hdi_diff = hdi_2019 - hdi_2010) %>%
  dplyr::mutate(
    hdi_diff_c = case_when(
      hdi_diff > 0 ~ "Higher HDI",
      hdi_diff == 0 ~ "No change",
      hdi_diff < 0 ~ "Lower HDI",
      is.na(hdi_diff) ~ "No Data"
    )
  ) %>%
  dplyr::mutate(gdi_diff = gdi_2019 - gdi_2010) %>%
  dplyr::mutate(
    gdi_diff_c = case_when(
      gdi_diff > 0 ~ "Higher GDI",
      gdi_diff == 0 ~ "No change",
      gdi_diff < 0 ~ "Lower GDI",
      is.na(gdi_diff) ~ "No Data"
    )
  ) %>%
  dplyr::mutate(
    gdi_gii = case_when(
      (gdi_diff > 0 & 
         gii_diff < 0) ~ "Better GDI and GII",
      (gdi_diff <= 0 &
         gii_diff >= 0) ~ "Worsened GDI and GII",
      (gdi_diff > 0 &
         gii_diff >= 0) ~ "Better GDI BUT worsened GII",
      (gdi_diff <= 0 & 
         gii_diff < 0) ~ "Worsened GDI BUT better GII",
      is.na(gdi_diff) | 
        is.na(gii_diff) ~ "No Data"
    )
  ) %>%
  dplyr::mutate(gdi_gii = factor(
    gdi_gii,
    levels = c(
      "Better GDI and GII",
      "Worsened GDI and GII",
      "Better GDI BUT worsened GII",
      "Worsened GDI BUT better GII",
      "No Data"
    )
  )
)
```

------------------------------------------------------------------------

### Plot gdi_gii in the Static Map

```{r}
tmap_mode("plot")
tmap_mode("view")
```

```{r}
install.packages("providers")
library(providers)
```

```{r}
tm_gdi_gii <- tm_shape(world) +
  tm_polygons(
    "gdi_gii",
    title = "Change from 2010 to 2019",
    border.col = "lightgrey",
    palette = c("grey", "darkgrey", "darkred", "red", "white")
  ) +
  
  tm_legend(show = TRUE) +
  tm_credits(text = "Shaun Hoang | Source: UN Data",
             size = 0.5,
             position = c(0.75, 0.05)) +
  tm_credits(
    text = "GDI: Gender Development Index\nGII: Gender Inequality Index",
    size = 0.5,
    position = c(0.03, 0.92),
    fontface = "italic",
    bg.color = "white",
    bg.alpha = 0.5
  ) +
  tm_layout(
    main.title = "Battle of the indices: GDI vs GII to measure gender inequalities",
    main.title.size = 0.8,
    main.title.fontface = "bold",
    main.title.position = c(0.03, 0.85),
    legend.title.size = 0.7,
    legend.position = c(0.03, 0.13),
    legend.text.size = 0.5,
    frame = FALSE
  )+
  tm_view(view.legend.position = c("left", "bottom"),
          control.position =  c("right", "bottom")
          ) 
tm_gdi_gii


```

```{r}
tmap_save(tm_gdi_gii,)
```
